---
title: "ST558 Project 2"
author: "Mu-Tien, Lee"
output: 
  github_document:
      toc: true
params:
  w: 1
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, error = FALSE, message=FALSE)
```

# Require package
```{r require package }
#install.packages("")
library(knitr)
library(rmarkdown)
library(MuMIn)
library(tidyverse)
library(caret)
library(corrplot)
library(readxl)
library(caret)
library(ggiraphExtra)
library(knitr)
library(ggplot2)
library(ggpubr)
library(rpart.plot)
library(rpart)
library(DT)
```

# Read in data
```{r data}

#read in hour data
HourData <- read.csv("hour.csv")
HourData<- HourData %>% select(-casual, -registered)
HourData$yr <- as.factor(HourData$yr)
HourData$holiday <- as.factor(HourData$holiday)
HourData$workingday <- as.factor(HourData$workingday)

#filter data by weekday
HourData <-HourData %>% filter(weekday==params$w)
#showing data
HourData <-HourData %>% select(-weekday, -workingday,-instant)
tbl_df(HourData)

#Separate dataset into train (70%) and test (30%) data set
set.seed(1997)
train <- sample(1:nrow(HourData), size = nrow(HourData)*0.7)
test <- dplyr::setdiff(1:nrow(HourData), train)
HourDataTrain <- HourData[train, ]
HourDataTest <- HourData[test, ]


```

# Summarize the training data
```{r summarizing data}
# plot the histogram of rental count
hist <- ggplot(data=HourDataTrain, aes(x=cnt))+geom_histogram(binwidth = 20, aes(color=yr))
hist <-hist+labs(title="Histogram of the retal count", x="rental count")
hist <-hist+scale_fill_discrete(labels=c(2011,2012))
hist

#prin out summary table for tempature humidity and windspeed
sum <- HourDataTrain%>% select(c(temp, atemp, hum, windspeed))
kable(apply(sum, 2,summary), caption="Numeric Summary for weather measurement", format ="html")

#plot the boxplot of tempature humidity and windspeed (not genralized amount)
#plot base
boxplot <- ggplot(data = HourDataTrain, aes(x=season))
#adding 4 variables
tem <-boxplot+geom_boxplot(aes(y=temp*41, group=season))+labs(y="Tempature (c)", title = "boxplot for weather measurement")
fetem <-boxplot+geom_boxplot(aes(y=atemp*50, group=season))+labs(y="Feeling Tempature (c)")
hum <-boxplot+geom_boxplot(aes(y=hum*100, group=season))+labs(y="Humidity")
wind <-boxplot+geom_boxplot(aes(y=windspeed*67, group=season))+labs(y="Wind Speed")
#combine 4 plots into 1
ggarrange(tem, fetem, hum , wind, ncol = 2, nrow = 2)

# plot the count distribution among time and weather
# by time
barplot1<-ggplot(data = HourDataTrain, aes(x=hr))+geom_col(aes(y=cnt, fill=yr))+facet_wrap(~mnth)
barplot1 <- barplot1+labs(x="time", y="Rental Count", title="Retal count distribution by month" )
barplot1+scale_fill_discrete(name="year", labels=c(2011,2012))

# by weather
barplot2 <-ggplot(data = HourDataTrain, aes(x=weathersit))+geom_col(aes(y=cnt, fill=yr))+facet_wrap(~mnth)
barplot2 <- barplot2+labs(x="Weather situation, 1: clear day, 2: misty day, 3:rain or snow", y="Rental Count", title="Retal count distribution by month" )
barplot2+scale_fill_discrete(name="year", labels=c(2011,2012))
```

# Training Model
Here I use two different method, First one useing a tree-based models using leave one out cross validation. Second, I use the boosted tree model with cross validation. Both two training are done using the `train` function from `caret` package. The data was cantered and scaled before training.Since our response variables is continuous response, I choose to use Regression tree.

## Tree-based model
```{r tree based model, cache=TRUE}
# set up training control, using leave one out cross validation.
set.seed(615)
trctrl <- trainControl(method = "LOOCV", number = 1)

# getModelInfo("rpart")
# training using regression tree models with cp in [0.001, 0.002]
# since the cp seems have to be really small when I used the default cp to train

model1 <- cnt~season+yr+mnth+hr+holiday+weathersit+temp+atemp+hum+windspeed

RegTree_fit1 <- train(model1, data = HourDataTrain, method = "rpart",
                 trControl=trctrl,
                 preProcess = c("center", "scale"),
                 tuneGrid=expand.grid(cp=seq(0.0001,0.0015,0.00005))
)

# show the training result
RegTree_fit1

# plot the RMSE of selected cp
plot(RegTree_fit1)

# plot my final tree model
rpart.plot(RegTree_fit1$finalModel)

```

## Boosted tree model

```{r boosting tree model, cache=TRUE, results="hide"}
# set up training control, using cross validation with 10 folder
set.seed(615)
trctrl <- trainControl(method = "cv", number = 10)

# training using boosted tree models with boosting interation in [200,400] and try max tree depth 5~9
model2 <- cnt~season+yr+mnth+hr+holiday+weathersit+temp+atemp+hum+windspeed
RegTree_fit2 <- train(model2, data = HourDataTrain, method = "gbm",
                trControl=trctrl,
                preProcess = c("center", "scale"),
                tuneGrid=expand.grid(n.trees=seq(800,1250,20),
                                     interaction.depth=6:10,
                                     shrinkage=0.1, n.minobsinnode=10)
                 )

```
```{r boosted result}

# show the training result
RegTree_fit2$bestTune

# plot the RMSE of selected parameters
plot(RegTree_fit2)

```

# Predicting using the best tree-base model 
```{r predicting tree model}

# predict use predict function
tree_pred <- predict(RegTree_fit1, newdata = HourDataTest)

#Calculate the Root MSE
RMSE_tree<- sqrt(mean((tree_pred-HourDataTest$cnt)^2))
label <- paste0("RMSE =", RMSE_tree)

# plot the prediction
count <- data.frame(true_count=HourDataTest$cnt,prediction=tree_pred )
predPlot <- ggplot(data=count, aes(x=true_count,y=prediction))
predPlot <- predPlot+labs(title="Prediction V.s. True Count using tree-base model")+geom_point()
predPlot <- predPlot+geom_smooth(color="orange")+geom_abline(aes(intercept=0,slope=1), color="blue")
predPlot <- predPlot+geom_text(x=200, y=800,label=label, color="brown")
predPlot

```

# Predicting using the best boosted-tree model 
```{r predicting boosted model}

# predict use predict function
boosted_pred <- predict(RegTree_fit2, newdata = HourDataTest)

#Calculate the Root MSE
RMSE_boosted <- sqrt(mean((boosted_pred-HourDataTest$cnt)^2))
lab <- paste0("RMSE =", RMSE_boosted)
# plot the prediction
count2 <- data.frame(True_count=HourDataTest$cnt,prediction=boosted_pred )
pred_plot <- ggplot(data=count2, aes(x=True_count,y=prediction))
pred_plot <- pred_plot+labs(title="Prediction V.s. True Count using boosted model")+geom_point()
pred_plot <- pred_plot+geom_smooth(color="orange")+geom_abline(aes(intercept=0,slope=1), color="blue")
pred_plot <- pred_plot+geom_text(x=200, y=800,label=lab, color=" brown")
pred_plot
```